---
title: "Backtest Evaluation: Adaptive Ensemble vs Base Models"
subtitle: "Evaluation of Linear Pooling Ensemble (Backtest Loop)"
author: "Pipeline Evaluation"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    code_folding: hide
    fig_width: 14
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 14,
  fig.height = 8
)

library(dplyr)
library(readr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(knitr)
library(DT)
library(scoringutils)
library(hubUtils)
library(hubData)

log_shift <- function(x, offset = 1) {
  log(x + offset)
}

summarise_relative <- function(scores, by_cols, metric, baseline) {
  by_rel <- setdiff(by_cols, "model")
  scores_with_skill <- scoringutils::add_relative_skill(
    scores,
    compare = "model",
    metric = metric,
    baseline = baseline,
    by = by_rel
  )
  summary <- scoringutils::summarise_scores(scores_with_skill, by = by_cols)
  scaled_col <- paste0(metric, "_scaled_relative_skill")
  if (scaled_col %in% names(summary)) {
    summary <- summary %>% rename(scaled_rel_skill = !!rlang::sym(scaled_col))
  }
  summary
}

render_table <- function(df, caption_text) {
  df <- as.data.frame(df)
  numeric_cols <- names(df)[vapply(df, is.numeric, logical(1))]
  float_cols <- numeric_cols[vapply(
    df[numeric_cols],
    function(x) {
      x_num <- suppressWarnings(as.numeric(x))
      x_num <- x_num[!is.na(x_num)]
      if (length(x_num) == 0) {
        return(FALSE)
      }
      any(abs(x_num - round(x_num)) > 1e-9)
    },
    logical(1)
  )]
  table <- DT::datatable(
    df,
    rownames = FALSE,
    caption = caption_text,
    options = list(pageLength = 10, autoWidth = TRUE)
  )
  if (length(float_cols) > 0) {
    table <- DT::formatRound(table, columns = float_cols, digits = 2)
  }
  table
}
```

Generated on: `r Sys.time()`

# Overview

This report evaluates the performance of the **Adaptive Ensemble** generated via the retrospective backtest loop (Linear Pooling) against the retrospective base models and the FluSight baseline using Reich Lab scoring conventions.

```{r load-data}
hub_path <- path.expand("~/My Drive/Research/Faculty/SantillanaLab/FluSight-forecast-hub")

if (dir.exists(file.path(getwd(), "forecasts")) && dir.exists(file.path(getwd(), "joint_twostage_pool_test"))) {
  project_root <- normalizePath(getwd())
} else if (dir.exists(file.path(getwd(), "..", "forecasts")) && dir.exists(file.path(getwd(), "..", "joint_twostage_pool_test"))) {
  project_root <- normalizePath(file.path(getwd(), ".."))
} else if (dir.exists(file.path(getwd(), "..", "..", "forecasts")) && dir.exists(file.path(getwd(), "..", "..", "joint_twostage_pool_test"))) {
  project_root <- normalizePath(file.path(getwd(), "..", ".."))
} else {
  stop("Could not infer project_root containing forecasts/ and joint_twostage_pool_test/")
}

prospective_dir <- file.path(project_root, "forecasts", "prospective")
ensemble_dir <- file.path(project_root, "forecasts", "retrospective", "ensemble")
ensemble_shrink_dir <- file.path(project_root, "forecasts", "retrospective", "ensemble_shrink")
ensemble_loc_dir <- file.path(project_root, "forecasts", "retrospective", "ensemble_loc")
ensemble_knn_dir <- file.path(project_root, "forecasts", "retrospective", "ensemble_knn")
ensemble_meta_dir <- file.path(project_root, "forecasts", "retrospective", "ensemble_meta")
ensemble_gate_dir <- file.path(project_root, "forecasts", "retrospective", "ensemble_gate")
ensemble_robust_dir <- file.path(project_root, "forecasts", "retrospective", "ensemble_robust")
ensemble_tier_dir <- file.path(project_root, "forecasts", "retrospective", "ensemble_tier")
joint_backtest_dir <- file.path(project_root, "joint_twostage_pool_test", "outputs")
season_start_date <- as.Date("2025-11-01")
recent_weeks <- 4
include_us <- FALSE
stopifnot(dir.exists(hub_path))
stopifnot(dir.exists(prospective_dir))

hub_con <- connect_hub(hub_path)
hub_forecasts <- hub_con |>
  dplyr::filter(
    output_type == "quantile",
    reference_date >= season_start_date,
    target == "wk inc flu hosp"
  ) |>
  dplyr::collect() |>
  as_model_out_tbl() |>
  dplyr::mutate(
    reference_date = as.Date(reference_date),
    target_end_date = as.Date(target_end_date)
  )

parse_horizon_from_file <- function(path) {
  file_name <- basename(path)
  match_pos <- regexpr("_h([1-4])", file_name, perl = TRUE)
  if (match_pos[1] == -1) {
    return(NA_integer_)
  }
  as.integer(substr(file_name, match_pos[1] + 2, match_pos[1] + 2))
}

parse_model_from_file <- function(path) {
  file_name <- basename(path)
  sub("_h[1-4]_prospective_[0-9]{8}\\.csv$", "", file_name)
}

load_local_forecasts <- function(model_id, dir_path, file_pattern, season_start) {
  empty <- tibble::tibble(
    reference_date = as.Date(character()),
    target = character(),
    horizon = integer(),
    target_end_date = as.Date(character()),
    location = character(),
    output_type = character(),
    output_type_id = character(),
    value = double(),
    model_id = character()
  )
  if (!dir.exists(dir_path)) {
    return(empty)
  }
  files <- list.files(dir_path, pattern = file_pattern, full.names = TRUE)
  if (length(files) == 0) {
    return(empty)
  }
  out <- lapply(files, function(file_path) {
    horizon_file <- parse_horizon_from_file(file_path)
    if (is.na(horizon_file)) {
      stop("Unable to parse horizon from filename: ", file_path)
    }
    df <- readr::read_csv(file_path, show_col_types = FALSE)
    df %>%
      filter(output_type == "quantile") %>%
      mutate(
        reference_date = as.Date(reference_date) + 7,
        target_end_date = as.Date(target_end_date),
        horizon = horizon_file - 1,
        location = as.character(location),
        location = ifelse(grepl("^[0-9]+$", location),
                          sprintf("%02d", as.integer(location)),
                          location),
        output_type_id = as.character(output_type_id),
        model_id = model_id
      ) %>%
      select(reference_date, target, horizon, target_end_date, location,
             output_type, output_type_id, value, model_id)
  })
  bind_rows(out) %>%
    filter(reference_date >= season_start)
}

load_local_forecasts_from_filename <- function(dir_path, file_pattern, season_start) {
  empty <- tibble::tibble(
    reference_date = as.Date(character()),
    target = character(),
    horizon = integer(),
    target_end_date = as.Date(character()),
    location = character(),
    output_type = character(),
    output_type_id = character(),
    value = double(),
    model_id = character()
  )
  if (!dir.exists(dir_path)) {
    return(empty)
  }
  files <- list.files(dir_path, pattern = file_pattern, full.names = TRUE)
  if (length(files) == 0) {
    return(empty)
  }
  out <- lapply(files, function(file_path) {
    horizon_file <- parse_horizon_from_file(file_path)
    if (is.na(horizon_file)) {
      stop("Unable to parse horizon from filename: ", file_path)
    }
    model_id <- parse_model_from_file(file_path)
    df <- readr::read_csv(file_path, show_col_types = FALSE)
    df %>%
      filter(output_type == "quantile") %>%
      mutate(
        reference_date = as.Date(reference_date) + 7,
        target_end_date = as.Date(target_end_date),
        horizon = horizon_file - 1,
        location = as.character(location),
        location = ifelse(grepl("^[0-9]+$", location),
                          sprintf("%02d", as.integer(location)),
                          location),
        output_type_id = as.character(output_type_id),
        model_id = model_id
      ) %>%
      select(reference_date, target, horizon, target_end_date, location,
             output_type, output_type_id, value, model_id)
  })
  bind_rows(out) %>%
    filter(reference_date >= season_start)
}

load_backtest_ensemble <- function(dir_path, file_pattern, season_start, model_id,
                                   reference_shift_days = 0, allowed_refs = NULL,
                                   latest_only = FALSE) {
  empty <- tibble::tibble(
    reference_date = as.Date(character()),
    target = character(),
    horizon = integer(),
    target_end_date = as.Date(character()),
    location = character(),
    output_type = character(),
    output_type_id = character(),
    value = double(),
    model_id = character()
  )
  if (!dir.exists(dir_path)) {
    return(empty)
  }
  files <- list.files(dir_path, pattern = file_pattern, full.names = TRUE)
  if (length(files) == 0) {
    return(empty)
  }
  if (latest_only && length(files) > 1) {
    info <- file.info(files)
    files <- files[order(info$mtime, decreasing = TRUE)][1]
    message(sprintf(
      "Using latest file for %s: %s",
      model_id, basename(files)
    ))
  }
  out <- lapply(files, function(file_path) {
    df <- readr::read_csv(file_path, show_col_types = FALSE)
    df %>%
      filter(output_type == "quantile") %>%
      mutate(
        reference_date = as.Date(reference_date) + reference_shift_days,
        target_end_date = as.Date(target_end_date),
        horizon = as.integer(horizon),
        location = as.character(location),
        location = ifelse(grepl("^[0-9]+$", location),
                          sprintf("%02d", as.integer(location)),
                          location),
        output_type_id = as.character(output_type_id),
        model_id = model_id
      ) %>%
      select(reference_date, target, horizon, target_end_date, location,
             output_type, output_type_id, value, model_id)
  })
  result <- bind_rows(out) %>%
    filter(reference_date >= season_start)
  if (!is.null(allowed_refs)) {
    result <- result %>% filter(reference_date %in% allowed_refs)
  }
  result
}

local_forecasts <- bind_rows(
  load_local_forecasts(
    model_id = "SVM",
    dir_path = prospective_dir,
    file_pattern = "^SVM_h[1-4]_prospective_[0-9]{8}\\.csv$",
    season_start = season_start_date
  ),
  load_local_forecasts(
    model_id = "ARIMA",
    dir_path = prospective_dir,
    file_pattern = "^ARIMA_h[1-4]_prospective_[0-9]{8}\\.csv$",
    season_start = season_start_date
  ),
  load_local_forecasts(
    model_id = "LGBM_blended",
    dir_path = prospective_dir,
    file_pattern = "^LGBM-blended_h[1-4]_prospective_[0-9]{8}\\.csv$",
    season_start = season_start_date
  ),
  load_local_forecasts_from_filename(
    dir_path = prospective_dir,
    file_pattern = "^TwoStage-FrozenMu-bounded.*_h[1-4]_prospective_[0-9]{8}\\.csv$",
    season_start = season_start_date
  )
) %>%
  distinct(model_id, reference_date, target, target_end_date, location, horizon,
           output_type_id, .keep_all = TRUE)

hub_keys <- hub_forecasts %>%
  mutate(
    reference_date = as.Date(reference_date),
    target_end_date = as.Date(target_end_date),
    horizon = as.integer(horizon),
    location = as.character(location),
    location = ifelse(grepl("^[0-9]+$", location),
                      sprintf("%02d", as.integer(location)),
                      location)
  ) %>%
  select(reference_date, horizon, target_end_date, location) %>%
  distinct()

allowed_backtest_refs <- sort(unique(hub_keys$reference_date))

adaptive_ensemble_forecasts <- load_backtest_ensemble(
  dir_path = prospective_dir,
  file_pattern = "^AdaptiveEnsemble_prospective_[0-9]{8}\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble",
  reference_shift_days = 0,
  allowed_refs = NULL
)

adaptive_backtest_forecasts <- load_backtest_ensemble(
  dir_path = ensemble_dir,
  file_pattern = "^AdaptiveEnsemble_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-backtest",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

adaptive_shrink_forecasts <- load_backtest_ensemble(
  dir_path = ensemble_shrink_dir,
  file_pattern = "^AdaptiveEnsemble_shrink_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-shrink",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

adaptive_loc_forecasts <- load_backtest_ensemble(
  dir_path = ensemble_loc_dir,
  file_pattern = "^AdaptiveEnsemble_loc_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-loc",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

adaptive_knn_forecasts <- load_backtest_ensemble(
  dir_path = ensemble_knn_dir,
  file_pattern = "^AdaptiveEnsemble_knn_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-knn",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

adaptive_meta_forecasts <- load_backtest_ensemble(
  dir_path = ensemble_meta_dir,
  file_pattern = "^AdaptiveEnsemble_meta_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-meta",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

adaptive_gate_forecasts <- load_backtest_ensemble(
  dir_path = ensemble_gate_dir,
  file_pattern = "^AdaptiveEnsemble_gate_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-gate",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

adaptive_robust_forecasts <- load_backtest_ensemble(
  dir_path = ensemble_robust_dir,
  file_pattern = "^AdaptiveEnsemble_robust_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-robust",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

adaptive_tier_forecasts <- load_backtest_ensemble(
  dir_path = ensemble_tier_dir,
  file_pattern = "^AdaptiveEnsemble_tier_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-tier",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

ensemble_hedge_forecasts <- load_backtest_ensemble(
  dir_path = file.path(project_root, "forecasts", "retrospective", "ensemble_hedge"),
  file_pattern = "^AdaptiveEnsemble_hedge_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-hedge",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

ensemble_hedge_test_forecasts <- load_backtest_ensemble(
  dir_path = file.path(project_root, "joint_twostage_pool_test", "outputs", "ensemble_hedge_retrospective"),
  file_pattern = "^AdaptiveEnsemble_hedge_retrospective_.*\\.csv$",
  season_start = season_start_date,
  model_id = "AdaptiveEnsemble-hedge-test",
  reference_shift_days = 7,
  allowed_refs = allowed_backtest_refs
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

joint_twostage_backtest_forecasts <- load_backtest_ensemble(
  dir_path = joint_backtest_dir,
  file_pattern = "^backtest_cov.*\\.csv$",
  season_start = season_start_date,
  model_id = "JointTwoStagePool-backtest",
  reference_shift_days = 0,
  allowed_refs = allowed_backtest_refs,
  latest_only = TRUE
) %>%
  inner_join(hub_keys, by = c("reference_date", "horizon", "target_end_date", "location"))

raw_forecasts <- bind_rows(
  hub_forecasts,
  local_forecasts,
  adaptive_ensemble_forecasts,
  adaptive_backtest_forecasts,
  adaptive_shrink_forecasts,
  adaptive_loc_forecasts,
  adaptive_knn_forecasts,
  adaptive_meta_forecasts,
  adaptive_gate_forecasts,
  adaptive_robust_forecasts,
  adaptive_tier_forecasts,
  ensemble_hedge_forecasts,
  ensemble_hedge_test_forecasts,
  joint_twostage_backtest_forecasts
)

raw_truth <- read_csv(file.path(hub_path, "target-data", "target-hospital-admissions.csv"),
                      show_col_types = FALSE) |>
  dplyr::mutate(date = as.Date(date))

```

# Scoring (Reich Lab)

```{r scoring}
raw_data <- raw_forecasts |>
  dplyr::filter(horizon > -1) |>
  dplyr::left_join(
    raw_truth |>
      dplyr::select(target_end_date = date, location, location_name, true_value = value),
    by = c("location", "target_end_date")
  ) |>
  dplyr::filter(!is.na(true_value)) |>
  dplyr::rename(model = model_id, quantile = output_type_id, prediction = value) |>
  dplyr::mutate(quantile = as.numeric(quantile))

raw_forecast <- scoringutils::as_forecast_quantile(
  raw_data,
  forecast_unit = c("model", "reference_date", "target_end_date", "location", "horizon"),
  observed = "true_value",
  predicted = "prediction",
  quantile_level = "quantile"
)
raw_scores <- scoringutils::score(raw_forecast)

wis_metric <- if ("interval_score" %in% names(raw_scores)) "interval_score" else "wis"

filter_scores <- function(df) {
  if (!include_us) {
    df <- df %>% filter(location != "US")
  }
  df
}

last_eval_sat <- max(as.Date(raw_scores$reference_date), na.rm = TRUE)
first_eval_sat_recent <- last_eval_sat - weeks(recent_weeks)
first_eval_sat_season <- season_start_date

recent_scores <- raw_scores %>%
  filter(reference_date >= first_eval_sat_recent, reference_date <= last_eval_sat) %>%
  filter_scores()

stopifnot("FluSight-baseline" %in% recent_scores$model)

recent_per_horizon <- summarise_relative(
  recent_scores,
  by_cols = c("model", "horizon"),
  metric = wis_metric,
  baseline = "FluSight-baseline"
) %>%
  left_join(
    recent_scores %>%
      group_by(model, horizon) %>%
      summarise(n_units = sum(!is.na(.data[[wis_metric]])), .groups = "drop"),
    by = c("model", "horizon")
  ) %>%
  rename(relative_wis = scaled_rel_skill) %>%
  arrange(horizon, is.na(relative_wis), relative_wis)

recent_overall <- summarise_relative(
  recent_scores,
  by_cols = c("model"),
  metric = wis_metric,
  baseline = "FluSight-baseline"
) %>%
  left_join(
    recent_scores %>%
      group_by(model) %>%
      summarise(n_units = sum(!is.na(.data[[wis_metric]])), .groups = "drop"),
    by = "model"
  ) %>%
  rename(relative_wis = scaled_rel_skill) %>%
  arrange(is.na(relative_wis), relative_wis)

season_scores <- raw_scores %>%
  filter(reference_date >= first_eval_sat_season, reference_date <= last_eval_sat) %>%
  filter_scores()

season_per_horizon <- summarise_relative(
  season_scores,
  by_cols = c("model", "horizon"),
  metric = wis_metric,
  baseline = "FluSight-baseline"
) %>%
  left_join(
    season_scores %>%
      group_by(model, horizon) %>%
      summarise(n_units = sum(!is.na(.data[[wis_metric]])), .groups = "drop"),
    by = c("model", "horizon")
  ) %>%
  rename(relative_wis = scaled_rel_skill) %>%
  arrange(horizon, is.na(relative_wis), relative_wis)

season_overall <- summarise_relative(
  season_scores,
  by_cols = c("model"),
  metric = wis_metric,
  baseline = "FluSight-baseline"
) %>%
  left_join(
    season_scores %>%
      group_by(model) %>%
      summarise(n_units = sum(!is.na(.data[[wis_metric]])), .groups = "drop"),
    by = "model"
  ) %>%
  rename(relative_wis = scaled_rel_skill) %>%
  arrange(is.na(relative_wis), relative_wis)
```

## Recent Per-Horizon Results

Recent = reference dates in the last 4 weeks of available submissions: from `r first_eval_sat_recent` through `r last_eval_sat`.

```{r recent-per-horizon-table}
render_table(recent_per_horizon, "Recent Per-Horizon Relative WIS (Reich Lab scoring)")
```

## Recent Combined Results

```{r recent-combined-table}
render_table(recent_overall, "Recent Combined Relative WIS (Reich Lab scoring)")
```

## Season Per-Horizon Results

```{r season-per-horizon-table}
render_table(season_per_horizon, "Season Per-Horizon Relative WIS (Reich Lab scoring)")
```

## Season Combined Results

```{r season-combined-table}
render_table(season_overall, "Season Combined Relative WIS (Reich Lab scoring)")
```
