---
title: "Backtest Evaluation: Adaptive Ensemble vs Base Models"
subtitle: "Evaluation of Linear Pooling Ensemble (Backtest Loop)"
author: "Pipeline Evaluation"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    code_folding: hide
    fig_width: 14
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 14,
  fig.height = 8
)

library(dplyr)
library(readr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(knitr)
library(DT)
```

Generated on: `r Sys.time()`

# Overview

This report evaluates the performance of the **Adaptive Ensemble** generated via the retrospective backtest loop (Linear Pooling) against the retrospective base models and the FluSight baseline.

**Evaluation Period:** From `2024-10-01` to present (based on available backtest files).

```{r load-data}
# 1. Load Actuals (Truth Data)
latest_file <- function(dir, pattern) {
  files <- list.files(dir, pattern = pattern, full.names = TRUE)
  if (length(files) == 0) return(NA_character_)
  files[order(files)][length(files)]
}

# Determine root path prefix based on where script is running
if (dir.exists('../data/imputed_sets')) {
  root_prefix <- '..'
} else if (dir.exists('data/imputed_sets')) {
  root_prefix <- '.'
} else {
  stop("Could not locate data/imputed_sets/ directory relative to current WD: ", getwd())
}

data_dir <- file.path(root_prefix, 'data/imputed_sets')
actual_path <- latest_file(data_dir, '^imputed_and_stitched_hosp_.*\\.csv$')

if(is.na(actual_path)) stop("No stitched data found in ", data_dir)
actual_raw <- read_csv(actual_path, show_col_types = FALSE)

# Map location_name -> FIPS
location_to_fips <- c(
  'Alabama'='01','Alaska'='02','Arizona'='04','Arkansas'='05','California'='06','Colorado'='08','Connecticut'='09','Delaware'='10',
  'District of Columbia'='11','Florida'='12','Georgia'='13','Hawaii'='15','Idaho'='16','Illinois'='17','Indiana'='18','Iowa'='19',
  'Kansas'='20','Kentucky'='21','Louisiana'='22','Maine'='23','Maryland'='24','Massachusetts'='25','Michigan'='26','Minnesota'='27',
  'Mississippi'='28','Missouri'='29','Montana'='30','Nebraska'='31','Nevada'='32','New Hampshire'='33','New Jersey'='34','New Mexico'='35',
  'New York'='36','North Carolina'='37','North Dakota'='38','Ohio'='39','Oklahoma'='40','Oregon'='41','Pennsylvania'='42','Puerto Rico'='72',
  'Rhode Island'='44','South Carolina'='45','South Dakota'='46','Tennessee'='47','Texas'='48','Utah'='49','Vermont'='50','Virginia'='51',
  'Washington'='53','West Virginia'='54','Wisconsin'='55','Wyoming'='56','US'='US'
)

actual_data <- actual_raw %>%
  transmute(location = location_to_fips[location_name], date = as.Date(date), actual_value = total_hosp) %>%
  filter(!is.na(location), !is.na(actual_value))

# 2. Load Forecasts

# Helper to load standard CSVs
load_forecast_csv <- function(path, model_name, horizon=NULL) {
  if (!file.exists(path)) return(NULL)
  df <- read_csv(path, show_col_types = FALSE) %>%
    mutate(
      model = model_name,
      reference_date = as.Date(reference_date),
      target_end_date = as.Date(target_end_date),
      location = as.character(location),
      output_type_id = as.numeric(output_type_id),
      value = as.numeric(value)
    ) %>%
    filter(output_type == 'quantile')
  
  # Standardize horizon if missing or inconsistent
  if (!is.null(horizon)) {
    # Horizon passed explicitly - use it directly (already 1-based)
    df$horizon <- horizon
  } else if (!'horizon' %in% names(df)) {
    # Try to infer from filename or data structure (fallback)
    # Ideally should be passed in.
    df$horizon <- NA_integer_
  } else {
    # Horizon exists in file - check if 0-based and convert
    # Only convert if max horizon is 3 (indicating 0-based: 0,1,2,3)
    if (max(df$horizon, na.rm = TRUE) == 3) {
      df$horizon <- df$horizon + 1
    }
  }
  
  df %>%
    select(model, reference_date, horizon, target_end_date, location, output_type_id, value)
}

all_forecasts_list <- list()

# A. Load Base Models (Retrospective)
# These are stored by horizon
for (h in 1:4) {
  # Baseline
  path <- sprintf('%s/forecasts/retrospective/Flusight-baseline_h%d_forecasts.csv', root_prefix, h)
  all_forecasts_list[[paste0('Base_h',h)]] <- load_forecast_csv(path, 'FluSight-baseline', h)
  
  # ARIMA
  path <- sprintf('%s/forecasts/retrospective/arima/ARIMA_h%d_forecasts.csv', root_prefix, h)
  all_forecasts_list[[paste0('ARIMA_h',h)]] <- load_forecast_csv(path, 'ARIMA', h)
  
  # LGBM t100
  path <- sprintf('%s/forecasts/retrospective/lgbm_enhanced_t100/TwoStage-FrozenMu_h%d_forecasts.csv', root_prefix, h)
  all_forecasts_list[[paste0('LGBMt100_h',h)]] <- load_forecast_csv(path, 'LGBM-t100', h)

  # LGBM t10
  path <- sprintf('%s/forecasts/retrospective/lgbm_enhanced_t10/TwoStage-FrozenMu_h%d_forecasts.csv', root_prefix, h)
  all_forecasts_list[[paste0('LGBMt10_h',h)]] <- load_forecast_csv(path, 'LGBM-t10', h)

  # LGBM bounded
  path <- sprintf('%s/forecasts/retrospective/lgbm_enhanced_t10_bounded/TwoStage-FrozenMu_h%d_forecasts.csv', root_prefix, h)
  all_forecasts_list[[paste0('LGBMbounded_h',h)]] <- load_forecast_csv(path, 'LGBM-bounded', h)

  # SVM
  # SVM files are named svm_retrospective_h{h}.csv
  path <- sprintf('%s/forecasts/retrospective/svm_t100/svm_retrospective_h%d.csv', root_prefix, h)
  all_forecasts_list[[paste0('SVM_h',h)]] <- load_forecast_csv(path, 'SVM', h)
}

# B. Load Adaptive Ensemble Backtest (Linear Pooling)
# These files are one per DATE, containing ALL horizons.
# forecasts/retrospective/ensemble/AdaptiveEnsemble_retrospective_YYYYMMDD.csv
ens_files <- list.files(file.path(root_prefix, 'forecasts/retrospective/ensemble'), pattern = 'AdaptiveEnsemble_retrospective_.*\\.csv', full.names = TRUE)

if (length(ens_files) > 0) {
  ens_list <- lapply(ens_files, function(fp) load_forecast_csv(fp, 'AdaptiveEnsemble'))
  ens_df <- bind_rows(ens_list)
  # Ensure horizon is handled (AdaptiveEnsemble outputs usually have 0-based horizons in column)
  # load_forecast_csv attempts to fix this, but let's double check
  # If max horizon is 3, it's 0-based.
  if (max(ens_df$horizon, na.rm=TRUE) == 3) {
     ens_df$horizon <- ens_df$horizon + 1
  }
  all_forecasts_list[['Ensemble']] <- ens_df
}

# Combine all
forecasts_all <- bind_rows(all_forecasts_list) %>%
  filter(!is.na(value))

```

# Scoring (Weighted Interval Score)

```{r scoring}
# WIS function (log scale)
calculate_wis_single <- function(qvals, qlevels, actual_value){
  qv <- as.numeric(qvals)
  ql <- as.numeric(qlevels)
  # Log scale transform for score calculation (CDC methodology)
  alog <- log(as.numeric(actual_value) + 1)
  
  if (!is.finite(alog)) return(NA_real_)
  
  keep <- is.finite(qv) & is.finite(ql)
  qv <- qv[keep]; ql <- ql[keep]
  if (length(qv) == 0 || length(ql) == 0) return(NA_real_)
  
  alphas <- unique(c(ql[ql <= 0.5], 1 - ql[ql > 0.5]))
  alphas <- alphas[is.finite(alphas) & alphas > 0]
  if (length(alphas) == 0) return(NA_real_)
  
  # Log transform forecasts
  qlog <- log(pmax(qv, 0) + 1)
  
  scores <- sapply(alphas, function(a){
    li <- suppressWarnings(which.min(abs(ql - a)))
    ui <- suppressWarnings(which.min(abs(ql - (1 - a))))
    if (length(li) == 0 || length(ui) == 0 || is.na(li) || is.na(ui)) return(NA_real_)
    L <- qlog[li]; U <- qlog[ui]
    if (!is.finite(L) || !is.finite(U)) return(NA_real_)
    
    width <- U - L
    pen <- if (is.finite(alog) && alog < L) (2/a) * (L - alog) else if (is.finite(alog) && alog > U) (2/a) * (alog - U) else 0
    if (is.finite(width) && is.finite(pen)) width + pen else NA_real_
  })
  
  if (all(is.na(scores))) NA_real_ else mean(scores, na.rm = TRUE)
}

# Calculate Scores
# Group by unique forecast instance (Model, Horizon, Date, Location)
scores_df <- forecasts_all %>%
  group_by(model, horizon, reference_date, target_end_date, location) %>%
  summarise(
    quantile_values = list(value), 
    quantile_levels = list(output_type_id), 
    .groups = 'drop'
  ) %>%
  inner_join(actual_data, by = c('target_end_date' = 'date', 'location')) %>%
  rowwise() %>%
  mutate(wis = calculate_wis_single(unlist(quantile_values), unlist(quantile_levels), actual_value)) %>%
  ungroup() %>%
  filter(!is.na(wis)) %>%
  # Exclude US and PR for aggregate summaries usually
  filter(!location %in% c('US', '72')) 

# Restrict to Common Set for Fair Comparison
# We want to compare models only on (Location, Date, Horizon) tuples where ALL models are present.
# Specifically, we care about comparing AdaptiveEnsemble vs Baseline vs Best Base Model.

# 1. Identify common tuples between AdaptiveEnsemble and Baseline
# (Base models might have different availability, but usually Baseline matches Ensemble)
common_set <- scores_df %>%
  select(model, horizon, target_end_date, location) %>%
  distinct() %>%
  mutate(present = 1) %>%
  pivot_wider(names_from = model, values_from = present) %>%
  filter(!is.na(`AdaptiveEnsemble`) & !is.na(`FluSight-baseline`)) %>%
  select(horizon, target_end_date, location) # Unique keys

scores_common <- scores_df %>%
  inner_join(common_set, by = c("horizon", "target_end_date", "location"))

```

# Results

## 1. Geometric Mean WIS (Per Horizon)

Lower is better. "Relative WIS" is Ratio to Baseline ( < 1.0 is better than baseline).

```{r summary-horizon}
summary_horizon <- scores_common %>%
  group_by(horizon, model, location) %>%
  # 1. Mean WIS per location (arithmetic)
  summarise(mean_wis_loc = mean(wis), .groups='drop') %>%
  group_by(horizon, model) %>%
  # 2. Geometric Mean across locations
  summarise(geo_mean_wis = exp(mean(log(mean_wis_loc))), .groups='drop') %>%
  group_by(horizon) %>%
  mutate(
    baseline_score = geo_mean_wis[model == "FluSight-baseline"],
    rel_wis = geo_mean_wis / baseline_score
  ) %>%
  ungroup() %>%
  arrange(horizon, rel_wis)

datatable(summary_horizon %>% mutate(across(where(is.numeric), round, 3)), 
          options = list(pageLength = 20))
```

## 2. Geometric Mean WIS (Combined Horizons)

Aggregated across all 4 horizons.

```{r summary-combined}
summary_combined <- scores_common %>%
  group_by(model, location) %>%
  summarise(mean_wis_loc = mean(wis), .groups='drop') %>%
  group_by(model) %>%
  summarise(geo_mean_wis = exp(mean(log(mean_wis_loc))), .groups='drop') %>%
  mutate(
    baseline_score = geo_mean_wis[model == "FluSight-baseline"],
    rel_wis = geo_mean_wis / baseline_score
  ) %>%
  arrange(rel_wis)

datatable(summary_combined %>% mutate(across(where(is.numeric), round, 3)))
```

## 3. Performance Over Time

Average WIS (across locations) by target date.

```{r time-plot}
time_scores <- scores_common %>%
  group_by(target_end_date, model) %>%
  summarise(mean_wis = mean(wis), .groups='drop')

ggplot(time_scores, aes(x=target_end_date, y=mean_wis, color=model)) +
  geom_line(size=1) +
  geom_point() +
  scale_y_log10() +
  labs(title="Average WIS over Time (Log Scale)", y="Mean WIS", x="Target Date") +
  theme_minimal()
```

## 4. Head-to-Head: Ensemble vs Baseline

Percentage of (Location, Week, Horizon) instances where Ensemble beat Baseline.

```{r head-to-head}
comparison <- scores_common %>%
  select(horizon, target_end_date, location, model, wis) %>%
  filter(model %in% c("AdaptiveEnsemble", "FluSight-baseline")) %>%
  pivot_wider(names_from = model, values_from = wis, values_fn = mean) %>%
  mutate(
    Ensemble_Better = `AdaptiveEnsemble` < `FluSight-baseline`
  )

win_rate <- comparison %>%
  group_by(horizon) %>%
  summarise(
    n = n(),
    win_pct = mean(Ensemble_Better) * 100
  )

kable(win_rate, digits=1, caption="Ensemble Win Rate vs Baseline (%)")
```

